%%% Verze pro jednostranný tisk:
%\documentclass[11pt,a4paper]{report}
%\usepackage[top=25mm,bottom=25mm,right=25mm,left=30mm,head=12.5mm,foot=12.5mm]{geometry}
%\let\openright=\clearpage

%%% For two-sided printing:
\documentclass[11pt,a4paper,twoside,openright]{report}
\usepackage[top=25mm,bottom=25mm,right=25mm,left=30mm,head=12.5mm,foot=12.5mm]{geometry}
\let\openright=\cleardoublepage

%%% Charts
\usepackage{pgf-pie}
\usepackage{pgfplots}

%%% Definition of useful macros
\input{makra}

%%% DEFINICE ZÁKLADNÍCH PROMĚNNÝCH
\def\TypPrace{BP}                % bakalářská práce/bachelor thesis
%\def\TypPrace{DP}               % diplomová práce/master thesis
%\def\Jazyk{cze}                  % čeština/czech
%\def\Jazyk{slo}                 % slovenština/slovak
\def\Jazyk{eng}                 % angličtina/english

%%% Title of the thesis in the language used in the text (exact according to assignment)
\def\NazevPrace{Development of a documentation generating application for .NET libraries}

\input{makra.txt}

%%% Author's name - Firstname and Lastname
\def\AutorPrace{Denis Akopyan}

%%% Year of submission and month (verbally) - month YYYY
\def\DatumOdevzdani{měsíc RRRR}

%%% Supervisor: First name and surname with titles
\def\Vedouci{Ing. Jarmila Pavlíčková, Ph.D.}

%%% Consultant: First name and surname with titles
\def\Konzultant{}

%%% Study program
\def\StudijniProgram{Applied Informatics}

%%% Study program - specialization
\def\Specializace{}

%%% Field of study
\def\StudijniObor{Applied Informatics}

%%% Optional thanks (the supervisor, the consultant, the borrower of software, literature, etc.)
\def\Podekovani{%
Poděkování.
}

%%% Abstrakt (doporučený rozsah cca 150-250 slov; nejedná se o zadání práce)
\def\Abstrakt{%
Zaměřením dané bakalářské práce je výzkum a vývoj aplikace pro generování dokumentace .NET knihoven. Hlavním cílem bylo vytvoření nástroje pro .NET vývojáře, jež chtějí extrahovat dokumentaci zdrojového kódu a exportovat ji do požadovaného výstupu, a zároveň zachovat aplikaci rozšiřitelnou, multiplatformní a uživatelsky přístupnou.\

Teoretická část práce se zabývá plánováním vývoje a očekáváným výsledkem projektu.

Empirická část popisuje vývoj samotný a zaměřuje se na tvorbu oddělené projektové struktury.

Výsledkem bakalářské práce byla funkční, multiplatformní a open-source aplikace, která umožňuje generování dokumentace z .NET knihoven, a zároveň je rozšiřitelná prostřednictvím modulů.

Práce došla k závěru, že je prostor pro rozšíření projektu o další moduly a funkce. Zavedení sofistikovaného testování navíc zajistí vysoce kvalitní kód, přiláká přispěvatele a zvýší důvěru uživatelů.
}
\def\AbstraktEN{%
This thesis aimed to research and develop a documentation-generating application for .NET libraries. 
The main objective was to create a tool for .NET developers that want to extract source code documentation and export it into their desired format while keeping the application extensible, cross platform, and easy to use.

The theoretical part of the thesis covered the development planning and the expected result of the 
project.

The empirical part of the thesis describes the exact process of development. Designing a decoupled 
project structure had the most attention dedicated to it.

The result of this thesis project was a working cross-platform and open-source application, which can 
generate documentation from .NET libraries and is extensible via plugins.

The thesis concludes that there is a prospect to enhance the project with more plugins and features. 
Additionally, the introduction of sophisticated testing will ensure a high-quality codebase, attract 
contributors, and raise trust amongst users
}

%%% 3 až 5 klíčových slov (doporučeno)
\def\KlicovaSlova{vývoj, tvorba, aplikace, generování, dokumentace, dotnet, .NET, knihoven}
\def\KlicovaSlovaEN{development, application, generating, documentation, dotnet, .NET, libraries}

%%% Kody podle klasifikace JEL
\def\JEL{JEL1, JEL2, JEL3}


%%% Title page and various mandatory information pages
\begin{document}
\include{zacatek}

%%% A page with automatically generated content of the thesis
\setcounter{tocdepth}{2}
\tableofcontents

%%% List of figures in the thesis
\openright
\listoffigures

%%% List of tables in the thesis (optionally)
\clearpage
\listoftables

%%% List of abbreviations in the thesis (optionally)
\include{zkratky}

\pagestyle{fancy}
%%% The individual chapters of the thesis are stored in separate files for clarity
\include{introduction}
\include{kap01}
\include{kap02}
\include{kap03}
\include{kap04}
% \include{...}
% \include{...}
\include{summary}

%%% Bibliography
\include{literatura}

%%% Attachments to thesis, if any. Each attachment must be referenced at 
%%% least once in your own text. The appendices are numbered.
\part*{\Prilohy}
\appendix
\include{app01}
\include{app02}
% \include{...}
% \include{...}

\end{document}
